Canopy leaf area of an evergreen Eucalyptus woodland does not respond to elevated

### 2015-04-27 manuscript.Rmd
abstract minor fix

Leaf area, quantified by the leaf area index (*L*), is a crucial driver of forest productivity, water use, and energy balance. Because *L* responds to environmental drivers, it can represent an important feedback to climate change, but its responses to rising [$CO_2$] and water availability of unmanaged forests have been poorly quantified. We studied canopy leaf area dynamics for `r nummon` months in a native evergreen Eucalyptus woodland exposed to Free Air $CO_2$ Enrichment (the EucFACE experiment), in a temperate (hot summer) climate where water limitation is common. We hypothesized that, because of expected higher productivity and increased water-use efficiency, *L* should increase with elevated [$CO_2$]. We estimated *L* from diffuse canopy transmittance and canopy photos, and measured monthly leaf litter production. Contrary to expectation, *L* did not respond to elevated [$CO_2$] in this timeframe. We found that *L* varied between `r rangeLAI[1]` and `r rangeLAI[2]`, but did not show a clear phenological pattern. Instead, increased *L* following leaf flushing occurred a few weeks after heavy rainfall, and *L* decreased steadily during periods of low rainfall. Leaf litter production was correlated to changes in *L*, both during periods of decreasing *L* (when no leaf growth occurred), and during periods of increasing *L* (active shedding of old foliage when new leaf growth occurred). We discuss these results in the context of modelling phenology of evergreen woodlands in response to environmental drivers.

>>>>

Leaf area, quantified by the leaf area index (*L*), is a crucial driver of forest productivity, water use, and energy balance. Because *L* responds to environmental drivers, it can represent an important feedback to climate change, but its responses to rising [$CO_2$] and water availability of unmanaged forests have been poorly quantified. We studied canopy leaf area dynamics for `r nummon` months in a native evergreen *Eucalyptus* woodland exposed to Free Air $CO_2$ Enrichment (the EucFACE experiment), in a temperate (hot summer) climate where water limitation is common. We hypothesized that, because of expected higher productivity and increased water-use efficiency, *L* should increase with elevated [$CO_2$]. We estimated *L* from diffuse canopy transmittance, and measured monthly leaf litter production. Contrary to expectation, *L* did not respond to elevated [$CO_2$] in this timeframe. We found that *L* varied between `r rangeLAI[1]` and `r rangeLAI[2]`, but did not show a clear annual pattern. Instead, dynamics of *L* showed a quick increase after heavy rainfall, and a steady during periods of low rainfall. Leaf litter production was correlated to changes in *L*, both during periods of decreasing *L* (when no leaf growth occurred), and during periods of increasing *L* (active shedding of old foliage when new leaf growth occurred). Leaf lifespan, estimated from mean *L* and total annual litter production, was significantly increased under elevated [$CO_2$] (p = 0.051). We discuss these results in the context of modelling phenology of evergreen woodlands in response to environmental drivers.
